---
title: "The Human Activity Recognition Research"
author: "Manuel Cerda"
date: "12/24/2017"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# About the Analisis

The purpose of this analysis is to classify the manner how people perfom an excercise based on the _Human Ativity Recognition_ research. We will use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants that were asked to perform barbell lifts correctly and incorrectly in 5 different ways. 

One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it.


## About the Data

The data was collected from devices such as _Jawbone Up_, _Nike FuelBand_, and _Fitbit_. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks.

The training data for this project are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source:
http://groupware.les.inf.puc-rio.br/har 

```{r loadData, include=FALSE}
trainUrl <-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
trainFile <- "./data/pml-training.csv"
testFile  <- "./data/pml-testing.csv"
if (!file.exists("./data")) {
  dir.create("./data")
}
if (!file.exists(trainFile)) {
  download.file(trainUrl, destfile=trainFile, method="curl")
}
if (!file.exists(testFile)) {
  download.file(testUrl, destfile=testFile, method="curl")
}
```

## Exploratory Analysis

```{r exploreData, include=T, message=F, warning=F}
require(ggplot2)
require(corrplot)
require(rpart)
require(rpart.plot)
require(caret)
require(randomForest)
require(rattle)

set.seed(123)

# Already downloaded to save some time
train <- read.csv("./data/pml-training.csv")
test  <- read.csv("./data/pml-testing.csv")
dim(train)

# Cleaning the data
trainClasse <- train$classe
train <- train[, colSums(is.na(train)) == 0] 
train <- train[, !(grepl("^X|timestamp|window", names(train)))]
train <- train[, sapply(train, is.numeric)]
train$classe <- trainClasse
test  <- test[, colSums(is.na(test)) == 0]
test  <- test[, !(grepl("^X|timestamp|window", names(test)))]
test  <- test[, sapply(test, is.numeric)]
test[,"problem_id"] <- list(NULL)

dim(train)

# Desired clasifications
levels(train$classe)
```


## Training Model

```{r modeling, include=T}
# Using a Random Forest algorithm
rf <- randomForest(classe ~ ., data=train, ntree=100)
rf
```


## Test Prediction

```{r predictTest, include=T}
pred <- predict(rf, test)
pred
```


# Appendix: Plots

```{r classeHist, include=T, echo=F}
ggplot(data=train, aes(classe, fill=classe)) +
  xlab("Classe") + ylab("Frequency") + 
  labs(title="Number of Observations per Classe (Training Data)") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_bar()
```

```{r rfError, include=T, echo=F}
err_df <- data.frame(err_rate = rf$err.rate[,'OOB'], n_trees=1:rf$ntree)
ggplot(err_df, aes(x=n_trees, y=err_rate)) +
  xlab("Number of Trees") + ylab("Error Rate") + 
  labs(title="Accuracy of the Random Forest") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_line()
```

```{r treePlot, include=T, echo=F, warning=F}
tree <- rpart(classe ~ ., data=train, method="class")
prp(tree, cex=0.3, tweak=1.5, uniform=T, main="Classification Tree (Training Data)")
```